Revisiting divisible, injective and flabby abelian groups from a constructive point of view

### 2022-09-27 slides-fischbachau2022.tex
Final touches

{\usebackgroundtemplate{\begin{minipage}{\paperwidth}\vspace*{0.00cm}\hfill\mbox{\includegraphics[width=0.2\textwidth]{turing-machine}\;\;}\end{minipage}}
\begin{frame}{Exploring the realizability model}
  \medskip
  \medskip
  \medskip
  \begin{tabular}{@{\!\!\!\!\!\!\!}l@{\,}llp{1.8cm}}
    \toprule
    & Statement & classical? & realizable? \\
    \midrule
    \normalnumber{1} & Every number is prime or not prime. & \ccmark{}
    (trivially) & \ccmark \\
    \normalnumber{2} & After every number there is a prime. & \ccmark & \ccmark \\
    \normalnumber{3} & Every map $\NN \to \NN$ has a zero or not. & \ccmark{} (trivially) & \cxmark \\
    \normalnumber{4} & Every map $\NN \to \NN$ is computable. & \cxmark &
    \qswitch{4}{5}{\ccmark}\only<1-4>{\,} \visible<5->{(trivially)} \\
    \normalnumber{5} & Every map $\RR \to \RR$ is continuous. & \cxmark &
    \qswitch{5}{6}{\ccmark} \\
    \normalnumber{6} & Markov's principle holds. & \ccmark{} (trivially) &
    \qswitch{6}{7}{\ccmark{} (if MP)} \\
    \normalnumber{7} & Countable choice holds. & \ccmark &
    \qswitch{7}{8}{\ccmark{} (always!)} \\
    \normalnumber{8} & Heyting arithmetic is categorical. & \cxmark &
    \qswitch{8}{9}{\ccmark{} (if MP)} \\
    \bottomrule
  \end{tabular}
  \medskip

  \only<1>{\ \\\ \\\ \\\ }
  \only<2>{\expl{1}{There is a machine which determines of any given
  number whether it is prime or not. \\\ }}
  \only<3>{\expl{2}{There is a machine which, given a number~$n$, computes a
  prime larger than~$n$. \\\ \\\ }}
  \only<4>{\expl{3}{There is a machine which, given a machine
  computing a map~$f : \NN \to \NN$, determines whether~$f$ is constantly
  zero or not. \\\ }}
  \only<5>{\expl{4}{There is a machine which, given a machine
  computing a map~$f : \NN \to \NN$, outputs a machine
  computing~$f$. \\\ }}
  \only<6>{\ \\\ \\\ \\\ }
  \only<7>{\expl{6}{There is a machine which, given a machine
  computing a map~$f : \NN \to \NN$ and given the promise that it is \notnot
  the case that~$f$ has a zero, determines a zero of~$f$.}}
  \only<8->{\expl{7}{\justifying There is a machine which, given a machine
  computing for every~$x \in \NN$ some~$y \in A$ together with a realizer
  of~$\varphi(x,y)$, outputs a machine computing a suitable choice
  function~$\NN \to A$.}}
\end{frame}}

@@

    \visible<2->{``Every function $\NN \to \NN$ is computable'' is realized by
    \textsf{cat}.}
    \medskip

@@

    \visible<3->{``Every function $\NN \to \NN$ is computable'' is \emph{not}
    realized.}
    \medskip

@@

    \visible<4->{``Every function $\NN \to \NN$ has a zero or not'' is realized by
    infinite search.}
    \medskip

@@

    \justifying
    \visible<6->{``Every function $\RR \to \RR$ is continuous'' is realized if,
    in the physical world, only finitely many computational
    steps can be carried out in finite time and if it is possible to form
    tamper-free private communication channels.}
  \end{enumerate}
\end{frame}

@@

@@

  \textbf{Thm.} For every formula~$\varphi$ and set of formulas~$\Gamma$:
  \vspace*{-0.6em}
  \begin{enumerate}
    \item Minimally, $\neg\neg(\varphi^{\neg\neg}) \Rightarrow
    \varphi^{\neg\neg}$.
    \item Minimally, $\varphi^{\neg\neg} \Leftrightarrow \neg\neg\varphi$ in
    case that~$\varphi$ is geometric
    ($R{\top}{\bot}{\wedge}{\vee}{\exists}{\bigvee}$).
    \item Classical logic proves~$\varphi$ from~$\Gamma$
    iff minimal logic proves~$\varphi^{\neg\neg}$ from~$\Gamma^{\neg\neg}$.
  \end{enumerate}
  \pause
  \vspace*{-0.3em}

>>>>

{\usebackgroundtemplate{\begin{minipage}{\paperwidth}\vspace*{0.00cm}\hfill\mbox{\includegraphics[width=0.2\textwidth]{turing-machine}\;\;}\end{minipage}}
\begin{frame}{Exploring the realizability model}
  \medskip
  \medskip
  \medskip
  \begin{tabular}{@{\!\!\!\!\!\!}l@{\,}llp{1.8cm}}
    \toprule
    & statement & classical? & realizable? \\
    \midrule
    \normalnumber{1} & Every number is prime or not prime. & \ccmark{}
    (trivially) & \ccmark \\
    \normalnumber{2} & After every number there is a prime. & \ccmark & \ccmark \\
    \normalnumber{3} & Every map $\NN \to \NN$ has a zero or not. & \ccmark{} (trivially) & \cxmark \\
    \normalnumber{4} & Every map $\NN \to \NN$ is computable. & \cxmark &
    \qswitch{4}{5}{\ccmark}\only<1-4>{\,} \visible<5->{(trivially)} \\
    \normalnumber{5} & Every map $\RR \to \RR$ is continuous. & \cxmark &
    \qswitch{5}{6}{\ccmark} \\
    \normalnumber{6} & Markov's principle holds. & \ccmark{} (trivially) &
    \qswitch{6}{7}{\ccmark{} (if MP)} \\
    \normalnumber{7} & Countable choice holds. & \ccmark &
    \qswitch{7}{8}{\ccmark{} (always!)} \\
    \normalnumber{8} & Heyting arithmetic is categorical. & \cxmark &
    \qswitch{8}{9}{\ccmark{} (if MP)} \\
    \bottomrule
  \end{tabular}
  \medskip

  \only<1>{\ \\\ \\\ \\\ }
  \only<2>{\expl{1}{There is a machine which determines of any given
  number whether it is prime or not. \\\ }}
  \only<3>{\expl{2}{There is a machine which, given a number~$n$, computes a
  prime larger than~$n$. \\\ \\\ }}
  \only<4>{\expl{3}{There is a machine which, given a machine
  computing a map~$f : \NN \to \NN$, determines whether~$f$ has a
  zero or not. \\\ }}
  \only<5>{\expl{4}{There is a machine which, given a machine
  computing a map~$f : \NN \to \NN$, outputs a machine
  computing~$f$. \\\ }}
  \only<6>{\ \\\ \\\ \\\ }
  \only<7>{\expl{6}{There is a machine which, given a machine
  computing a map~$f : \NN \to \NN$ and given the promise that it is \notnot
  the case that~$f$ has a zero, determines a zero of~$f$.}}
  \only<8->{\expl{7}{\justifying There is a machine which, given a machine
  computing for every~$x \in \NN$ some~$y \in A$ together with a realizer
  of~$\varphi(x,y)$, outputs a machine computing a suitable choice
  function~$\NN \to A$.}}
\end{frame}}

@@

    \visible<2->{``Every map $\NN \to \NN$ is computable'' is realized by
    \textsf{cat}.}
    \medskip

@@

    \visible<3->{``Every map $\NN \to \NN$ is computable'' is \emph{not}
    realized.}
    \medskip

@@

    \visible<4->{\mbox{``Every map $\NN \to \NN$ has a zero or not'' is realized by
    infinite search.}\\[0em]}
    \medskip

@@

    \justifying
    \visible<6->{``Every map $\RR \to \RR$ is continuous'' is realized if,
    in the physical world, only finitely many computational
    steps can be carried out in finite time and if it is possible to form
    tamper-free private communication channels.}
  \end{enumerate}
\end{frame}

@@

\addtocounter{framenumber}{-1}
\begin{frame}[plain]\frametitle{A classical logic fairy tale}
  \vspace*{-0.5em}
  \begin{changemargin}{-1em}{-1em}
    \setlength{\columnsep}{15pt}
    \begin{multicols}{2}
      \justifying
      \scriptsize

      \textbf{Narrator.} Once upon a time, in a kingdom far, far away, the
      queen of the land and of all Möbius strips called for her royal
      philosopher.

      \textbf{Queen.} Philosopher! I ask you to carry out the following order.
      Get me the Philosopher's Stone, or alternatively find out how one could
      produce arbitrary amounts of gold with it!

      \textbf{Philosopher.} But my queen! I haven't studied anything useful!
      How could I fulfill this order?

      \textbf{Queen.} That is not my concern. I'll see you again tomorrow.
      Should you not accomplish the task, I will take your head off.

      \textbf{Narrator.} After a long and wakeful night the philosopher was
      called to the queen again.

      \textbf{Queen.} Tell me! What do you have to report?

      \textbf{Philosopher.} It was not easy and I needed to follow lots of
      obscure references, but finally I actually found out how to use the
      Philosopher's Stone to produce arbitrary amounts of gold. But only I can
      conduct this procedure, your royal highness.

      \textbf{Queen.}
      Alright. So be it.

      \textbf{Narrator.} And so years passed by, during which the philosopher
      imagined herself to be safe. The queen searched for the stone on her own,
      but as long as she hadn't found it, the philosopher didn't need to worry.
      Yet one day the impossible happened: The queen has found the stone! And
      promptly called for her philosopher.

      \textbf{Queen.} Philosopher, look! I have found the Philosopher's Stone!
      Now live up to your promise! \emph{[She hands over the stone.]}

      \textbf{Philosopher.} Thank you. \emph{[She inspects the stone.]} This is
      indeed the Philosopher's Stone. Many years ago you asked me to either
      acquire the Philosopher's Stone or find out how to produce arbitrary
      amounts of gold using it. Now it's my pleasure to present to you the
      Philosopher's Stone. \emph{[She returns the stone.]}
    \end{multicols}
  \end{changemargin}
\end{frame}

@@

  \textbf{Thm.} For every formula~$\varphi$ and set of formulas~$\Gamma$:
  \vspace*{-0.6em}
  \begin{enumerate}
    \item Minimally, $\neg\neg(\varphi^{\neg\neg}) \Rightarrow
    \varphi^{\neg\neg}$.
    \item Minimally, $\varphi^{\neg\neg} \Leftrightarrow \neg\neg\varphi$ in
    case that~$\varphi$ is geometric
    ($R{\top}{\bot}{\wedge}{\vee}{\exists}{\bigvee}$).
    \item If~$\Gamma$ entails~$\varphi$ classically,
    then~$\Gamma^{\neg\neg}$ entails~$\varphi^{\neg\neg}$ minimally.
  \end{enumerate}
  \pause
  \vspace*{-0.3em}
